---
title: Estimating the effect of Krasner's tenure as District Attorney on criminal charges
author: "Vitaly Lorman"
date: "3/25/2021"
output: 
  github_document:
  number_sections: true
bibliography: DAbib.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(knitr)
library(lubridate)
library(scales)
library(reshape2)
library(MatchIt)
library(optmatch)
library(causalsens)
library(knitr)
```



## Introduction

Elected in November 2017, Philadelphia District Attorney (DA) Larry Krasner ran on a platform whose stated top priority was ending mass incarceration in Philadelphia [@Krasner1]. A self-described progressive DA, Krasner's campaign promised, among other things, to stop prosecuting "insufficient and insignificant cases", to treat addiction as a public safety issue, and to work to stop pursuing extreme sentences [@Austen]. Running for re-election in 2021, Krasner's campaign proclaimed a 40% decrease in the Philadelphia county jail population, reductions in sentence length, an end to charging marijuana possession cases, diversion of other drug possession cases to treatment, and a shifting of resources to solving serious crimes, particularly homicides and gun violence [@Krasner2].

We carry out an observational study to determine the effect of Krasner's tenure and policies as Philadelphia DA on the number of charges for different types of offenses. Our primary source of data regarding arrests and charges comes from the Philadelphia DA's Office (DAO), and includes counts of arrests and charges by categories of offenses from January 1, 2014 to present. In measuring the outcomes from cases subject to the treatment (being charged by Krasner's office), we restrict our focus to the period from January 1, 2018 (the beginning of Krasner's term) to March 15, 2020 (prior to the start of the COVID-19 pandemic in Philadelphia). Our control group consists of cases charged by Krasner's predecessor, DA Seth Williams, between January 1, 2014 and June 29, 2017. (Williams resigned from office when he pleaded guilty to bribery on June 29, and we exclude cases charged during the six month term of interim DA Kelley B. Hodge from our analysis.)

Both prior to and over the course of Krasner's tenure, the crime rate across various offenses, measured both by incidents reported by Philadelphia and nearby county police agencies to the FBI's Uniform Crime Reporting (UCR) Program and by police arrests reported by the DA's office, has varied substantially. Since the DA's office typically only enters the criminal justice process after an arrest has been made, we control for arrest counts by date across a variety of offense categories. We argue that the most substantial potential covariates that effect charge counts do so only through their effect on the number of arrests across various categories and that consequently, by controlling for arrests, we produce estimates of the effect of Krasner's tenure as DA on the number of charges brough for various offense types.

Since the arrest counts vary between our control group (offenses charged by DA Williams) and treatment group (offenses charged by DA Krasner), we mitigate potential bias by employing matching as part of our preprocessing. Our results consist of estimates of effects (together with confidence intervals) of Krasner's tenure as DA on the number of charges across the following categories: violent, property, drugs, firearms, other, and uncategorized. We adopt these groupings from those reported by the Philadelphia DA's Office and describe a further refinement of which offenses comprise each category in our description of the data below. We state our assumptions, conduct sensitivity analyses, and conclude with a discussion of future work that could be carried out to improve our causal estimates.

## Background on the criminal justice process in Philadelphia

The DAO typically only gets involved in a case once an arrest has been made (by either the Philedelphia Police Department or another agency such that jurisdiction for prosecuting the case falls under Philadelphia County). At the first stage of this process, the DAO decides whether to bring charges and what charges to bring. Subsequently, bail is set by a Bail Magistrate, taking into account input from both the DA's office and the attorney for the defense. In some cases, instead of bringing charges, the DAO may decide to divert a case into one of various diversionary programs. If charges are brought, a case proceeds through the judicial system to an outcome (verdict, guilty plea, dismissal, or withdrawal). One of the DA's many roles in this part of the process is making plea offers. After an outcome, the case proceeds to sentencing. The judge imposes a sentence with recommendations from the DAO. A single case may have many charges, and distinct charges may have different outcomes. Different outcomes subject to sentencing may receive different sentences, and these sentences may be imposed concurrently or consecutively. 

In this document, we are interested in the part of the process that involves charges (which correspond to statutes in the Pennsylvania Criminal Law). The reason we are interested in charges is because how and whether a cases is charged is typically the first decision made by the DAO, and it occurs before many other actors (defense lawyers, judges, etc.) enter the process; thus, it is also the outcome for which we are able to study the effect of Krasner's DAO with the fewest potential confounders. 

The publicly availably data from the DAO that we will be using in our analysis counts only the most serious charge (subject to the most severe sentence) for each case. To help our analysis and interpretability, we group the possible charges into six groups based on the DAO's grouping of these charges on their website @DAO. These groups, and the kinds of charges that comprise them, are as follows:

* **Violent charges:** Homicide, Non-fatal shooting, Rape", Robbery with a gun, Robbery (other), Aggravated assault with a gun, Aggravated assault (other), Other assaults, Sexual assaults and other sex offenses
* **Property charges:** Residential burglary, Commercial burglary, Theft of motor vehicle tag,  Theft from person, Theft from auto, Retail theft", Theft, Auto theft, Fraud Theft of Services, Embezzlement
* **Drug charges:** Drug possession, Drug sales, DUI
* **Firearms charges:** Illegal firearms possession
* **Other charges:** Prostitution or sex work, Patronizing prostitutes, Threats of violence
* **Uncategorized charges:** All other charges brough by the DAO that do not fall into the categories above.


## The data

### Source of data

The data we use in our observational study is publicly available and was downloaded on March 22, 2021 from the District Attorney's Office GitHub repository @DAOdata. Specifically, we use the data frames consisting of counts of charges and of arrests by date, grouped by categories of offenses which are similar, but not identical to the FBI UCR Program categories of Type I and Type II crime. 

Our cleaning and preprocessing (described further in the cleaning.rmd file) included the following steps:

* Group offenses for charges and arrests into 6 groups (described further below): violent, property, drugs, firearms, other, and uncategorized by date. Calculate daily charge and arrest total in each of these categories.
* Subset the charge and arrest data frames on common values of date_value and merge them by date_value, keeping just the totals for each of the 6 categories. 
* Filter to include only dates prior to 2017-06-29 and after 2018-01-01. Filter out dates after 2020-03-15 (to exclude judicial actions after the COVID-19 pandemic began effecting the Philadelphia courts). Create a binary treatment variable, assigning it to be FALSE for dates prior to when Krasner took office and TRUE for dates after. This produces the data frame "charges_all"
* Reshape into a long data frame with columns date_value, type (arrest or chage), group (violent, property, drugs, firearms, other, or uncategorized) and the counts for each combination of the these. This data frame is called "charges_all_long" below.

We read in the resulting datasets.

```{r read}

charges_all<-read.csv("charges_all.csv", row.names=1)
charges_all_long<-read.csv("charges_all_long.csv", row.names=1)
charges_all$date_value<-as.Date(charges_all$date_value)
charges_all_long$date_value<-as.Date(charges_all_long$date_value)
```

## Observational study design

We set up our observational study such that a single unit consists of a day on which criminal charges in the Philadelphia Municipal Court or Common Pleas court could be brought between January 1, 2014 and June 29, 2017, as well as between January 1, 2018 and March 15, 2020.


### Treatment

Given a unit in our study--a day with with a collection of criminal cases that receive potential charges--we define the treatment to be having the cases that day charged by the DA's office under Krasner. The control is having the cases that day charged by the office of Krasner's predecessor, former DA Seth Williams.

Note that our treatment is defined somewhat narrowly, and in particular, the treatment does not include other actions or conditions during Krasner's tenure as head of the DAO, even those initiated by the DAO. Our treatment is strictly defined whether Krasner's office charges the cases on a given day. See the Assumptions section for further discussion of this.

Also note that we are not studying the effect of Krasner himself but the effect of the entire DAO under Krasner's leadership (as opposed to the control of the DAO under Williams' leadership). We view the treatment as binary (see the Discussion section for more on this).

### Outcomes

As discussed above, we group offenses into six groups (violent, property, drugs, firearms, other, and uncategorized). Our outcomes of interest are, for each offense type, the number of cases in which a charge of that type was brought as the most serious charge in that case. We measure outcomes as a difference in daily means between treatment and control group as well as a percent change relative to the control group daily mean.

## Covariates

The number of charges on a given day, by category, is caused by a number of factors: the number of criminal incidents in previous weeks, the number of arrests, policing practices such as gathering of evidence and writing reports, criminal law (local, state, and federal), criminal records of defendants, police department policies, and DA office policy and practice regarding charges. The DAO is the final actor in a complicated process.

For most of the factors that come prior to the DAO's role in the process, their effect on our outcomes of interest (mean daily charges by offense category) is indirect and goes through the number of arrests. For instance, policy department policies regarding whether to make arrests for a given type of offense and how actively to pursue enforcing it will only effect charges through the number of arrests made for that type of offense. Thus, our covariates consist of daily arrest totals by offense category. By conditioning on these, we address confounding by many of the factors mentioned above. 

### Assumptions

We assume stable unit treatment values (SUTVA); that is, we assume that the outcomes of one unit (the count of most serious charges in cases from one day by type) do not vary with the treatments assigned to other units. In other words, given a day, the charge counts from that day do not vary depending on whether the cases from a different day were by Krasner's office or Williams'. This assumption could be violated in the case of reoffenders if the past criminal record of a defendant had an effect on the sorts of charges that were brought in their case. We assume that the effect of such instances is negligible, though a data set that contained information on reoffenders would be necessary to test this assumption.

We also assume that there is only one form of the treatment. This assumption would be violated if during Krasner's tenure as DA, the DAO changed its policies or practices in charging certain kinds of cases. This would be particularly of concern in the first few months of Krasner's tenure as DA when big changes occured in the DAO (for instance, 31 staffers, including prosecutors, left the DAO in the first week of Krasner's tenure). We address this assumption further and conduct a sensitivity analysis in the Discussion section of this document.

We assume that unobserved confounders have negligible impact on our estimates; that is, we make an ignorability assumption that, controlling for observed covariates, treatment assignment is independent of the potential outcomes. Unobserved confounders that are of potential concern are ones that have a direct effect on charge counts rather than an indirect one through arrest counts. These could include quality of policy reports and evidence-gathering, prior criminal records of defendants, defendant demographics, and local, state, or federal policies regarding charges for certain types of offenses. The data available to us does not contain many of these. We address how robust our analysis is to their impact through sensitivity analyses in the Discussion section.

Finally, we assume there is no post-treatment bias; that is, the distribution of arrest counts does not vary with treatment assignment. Days charged during Krasner's tenure as DA appear to have a different distribution of arrests from days charged prior to Krasner's tenure as DA. However, we recall that our definition of treatment is narrow: we say that a unit (day) in our study has received the treatment if the cases on that day were charged by Krasner's office, and a unit has received the control if the cases were charged by Williams' office. In particular, our definition of treatment does not include any actions by the DAO or anyone else that effect arrests: for each individual unit, arrest counts are pre-treatment covariates and are uneffected by the treatment assignment. While the DAO could well have a effect on arrest counts by, for instance, instituting a policy of not charging certain kinds of offenses causing the police not to make arrests for those offenses, this kind of action of the DAO does not fall within our definition of treatment. Instead, we simply estimate the effect of Krasner's DAO on charging the arrests they are given. Any other effects of the DAO on charges are outside the scope of our study.

### Outline of analysis

Our method of analysis will be to use a linear regression to estimate the treatment effect on charge daily counts for each type regressing on daily arrest counts by type. We analyze each type of offense separately. In order to ensure that the arrest count covariates are similarly distributed between the treatment and control groups, we implement nearest-neighbor matching with calipers (further discussed below). Upon achieving our desired balance, we fit a linear model to the matched data to estimate the desired effects. 

## Summary of trends in the data

Before we conduct our analysis, we begin by presenting some summaries of our data. First, we show the counts for charges and arrests by type. For this chart, we use a version of the charges_all_long data frame in which the dates between 2017-06-29 and 2018-01-01 and the dates after 2020-03-25 have not been excluded.

```{r trends}
all_dates_long<-read.csv("charges_all_long_full.csv", row.names=1)
all_dates_long$date_value<-as.Date(all_dates_long$date_value)

ggplot(data=all_dates_long, aes(x=date_value, y=value, col=type))+
  geom_point(alpha=0.02)+
  geom_smooth()+
  facet_wrap(~group, scale="free_y")+
  labs(title="Daily counts of arrests and charges by offense type", x="Date", y="Count")
  
```


The above scatterplots and loess curves show trends in arrests and charges. We see that in four of the five categories, charges track pretty closely with arrests; the exception is uncategorized offenses--this may be a result of the fact that it is broadly defined and includes a variety of unrelated types of offenses that do not fit in the other groups. We note that both charge and arrest counts for violent, property, drug, and other offenses all saw a decrease over the course of our timeframe, and firearm charges and arrests increased over the same time free, with a steep increase beginning in 2020.


Next, we restrict our focus to the time periods of our control group (2014-01-01 to 2017-06-29) and our treatment group (2018-01-01 to 2020-03-15) and show the control and treatment means for arrests and charges by type, along with the percent in change from the control mean.

```{r group means}
charges_all_long$treatment<-charges_all_long$date_value>="2018-01-01"
means<-charges_all_long %>%
  group_by(treatment, type, group) %>%
  summarise(mean=mean(value)) %>%
  dcast(group+type~treatment, value.var="mean")

colnames(means)[3:4]<-c("control", "treatment")
means$diff_in_means<-means$treatment-means$control
means$percent_diff_in_means<-100*means$diff_in_means/means$control
##differences in means of charges
charge_diffs<-means %>%
  filter(type=="charge") %>%
  select("group", "diff_in_means", "percent_diff_in_means")

kable(charge_diffs, caption="Differences in daily means of charge counts")
```

The point estimates displayed in the table show that in each offense category except for firearms, there was a decrease in daily charge counts from Williams' to Krasner's tenure as DA. This was largest, a 56 percent decrease, in the category of other offenses (which includes prostitution and sex work related offenses, as well as threats of violence). In the category of firearms, we see a 24 percent increase in charges.

```{r}

ggplot(data=means, aes(x=group, y=percent_diff_in_means, fill=type))+
  geom_bar(position="dodge", stat="identity")
  
```

The barplot above shows the percent difference in means for charges and arrests from Williams' to Krasner's tenure as DA. We see that for drug, firearm, other, and property, and violent offenses, arrests and charges both decreased, and for firearms, arrests and charges both increased. For uncategorized offenses we see that arrests in charges moved in opposite directions.

## Matching

Our effect estimates will come from fitting linear models, controlling for arrest count covariates for each of our six types of offenses. Without any preprocessing to balance the arrest count covariates in the treatment and control groups, we might see a biased estimate due to different levels of arrest counts by offense type in the treatment and control groups. For instance, our scatteplots above show lower daily arrest counts for several offense types prior to 2018. Thus, before fitting our linear models, in this section we employ matching to account for imbalance of covariates in our treatment and control groups following [@HIKS]. First, here is a plot of the distributions of mean arrest counts by type, along with a QQ plot for each plot:

```{r balance plots}
arrest_count_sum<-charges_all_long %>%
  filter(type=="arrest") %>%
  group_by(group, treatment) %>%
  summarise(mean_per_day=mean(value))

ggplot(data=arrest_count_sum, aes(x=group, y=mean_per_day, fill=treatment))+
  geom_bar(position="dodge", stat="identity")

##Construct data with empirical quantiles for control and treatment groups

par(mfrow=c(2,3))
for (g in unique(charges_all_long$group))
{
  x<-filter(charges_all_long, type=="arrest", group==g, !treatment)$value
  y<-filter(charges_all_long, type=="arrest", group==g, treatment)$value
  qqplot(x, y, main=g,
  xlab="Control",
  ylab="Treatment",
  xlim=c(min(x, y), max(x, y)),
  ylim=c(min(x, y), max(x, y)))
  abline(0,1)
}
#mtext("QQ plots for mean daily arrest counts by offense group", outer=TRUE,side=3,line=0, cex=1.5)
```

We see imbalance for some of the arrest types, particularly property, drugs, and other.

To address this imbalance, we implementing matching on propensity scores. Following @HIKS and appealing to the "propensity score tautology", our sole concern is achieving balance on the covariates, and so whatever matching method we implement that yields our desired level of balance will do the job. We experimented with a variety of matching methods (further detailed in the "Matching.Rmd" document), and we observed a tradeoff between sample sizes of matched data and balance. For example, using subclassification matching with 6 classes, we were able to get good balance, but at the expense of one of the subclasses having as few as 19 control units (having started with 1275 control units). 

We settled on variable ratio nearest-neighbor matching with calipers of size 0.1 (standard deviations of the propensity scores within which to draw matches). Since the higher end of the propensity score distribution was overrepresented in the treatment group, the price we paid for our balance was about 31% of our treated observations going unmatched. The Discussion section contains further analysis of this methodological choice, and the supplementary "matching.Rmd" file contains summaries for the balance we achieved using this matching method, along with an exploration of other matching methods.

Here we print a few plots that demonstrate the covariate balance our matching achieves.

```{r calipers}
mcal1.out<-matchit(treatment~arrests_violent+arrests_property+arrests_drugs+arrests_other+arrests_firearms+arrests_uncategorized, data=charges_all, method="nearest", caliper=0.1, ratio=1275/804, min.controls=1, max.controls=3)
scal1.out <- summary(mcal1.out, standardize = TRUE)
#scal1.out
plot(scal1.out)
```

The Love plot above shows the standardized mean difference measure of balance for each covariate. The vertical dashed lines are drawn within .1 standardized mean difference units of zero, and we see that for each covariate, the standardized mean differences between treatment and control covariates all fall within this margin of zero. 

```{r}
plot(mcal1.out,  type = "jitter", interactive = FALSE)
plot(mcal1.out,  type = "hist")
plot(mcal1.out, type="qq")

matched_data2<-match.data(mcal1.out)
matched_data2$treatment<-as.numeric(matched_data2$treatment)
```

The jitter plot above shows that our matching method required dropping treatment units from the high end of the propensity score distribution and control units from the low end of the propensity score distrbution. The histogram shows very similar propensity score distributions between treatment and control groups in the matched data, and the QQ plots show that matching has resulted in similar distributions not just for propensity scores, but for arrest counts for each offense type. 

## Fitting linear models

Now that we have achieved balance between treatment and control groups, we fit a linear model to the data. For each outcome of interest (charge counts for each of the 6 offense groups), we regress on the binary treatment variable and the arrest counts for each of the 6 offense groups.

We summarize each linear model below.

```{r lin}
lm_violent<-lm(charges_violent~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data2)
#x.out <- setx(z.out, data = match.data(m.out), fn = NULL, cond = TRUE)
#s.out <- sim(z.out, x = x.out)
summary(lm_violent)

lm_property<-lm(charges_property~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data2)
#x.out <- setx(z.out, data = match.data(m.out), fn = NULL, cond = TRUE)
#s.out <- sim(z.out, x = x.out)
summary(lm_property)

lm_drugs<-lm(charges_drugs~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data2)
#x.out <- setx(z.out, data = match.data(m.out), fn = NULL, cond = TRUE)
#s.out <- sim(z.out, x = x.out)
summary(lm_drugs)


lm_firearms<-lm(charges_firearms~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data2)
#x.out <- setx(z.out, data = match.data(m.out), fn = NULL, cond = TRUE)
#s.out <- sim(z.out, x = x.out)
summary(lm_firearms)

lm_other<-lm(charges_other~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data2)
#x.out <- setx(z.out, data = match.data(m.out), fn = NULL, cond = TRUE)
#s.out <- sim(z.out, x = x.out)
summary(lm_other)

lm_uncategorized<-lm(charges_uncategorized~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data2)
#x.out <- setx(z.out, data = match.data(m.out), fn = NULL, cond = TRUE)
#s.out <- sim(z.out, x = x.out)
summary(lm_uncategorized)
```

## Results: summary of estimated effects

In the table below, we summarise our estimated effect for each offense group. We report the mean effect as well as the 95% confidence interval in both absolute counts and as percent change relative to the control mean.

```{r}
charge_type<-c("violent", "property", "drugs", "firearms", "other", "uncategorized")
effect_sum<-as.data.frame(charge_type)
effect_sum[,"treatment_effect"]<-
  rbind(lm_violent$coefficients["treatment"],
    lm_property$coefficients["treatment"],
    lm_drugs$coefficients["treatment"],
    lm_firearms$coefficients["treatment"],
    lm_other$coefficients["treatment"],
    lm_uncategorized$coefficients["treatment"])
                                
effect_sum<-cbind(effect_sum,
  rbind(confint(lm_violent)["treatment",],
    confint(lm_property)["treatment",],
    confint(lm_drugs)["treatment",],
    confint(lm_firearms)["treatment",],
    confint(lm_other)["treatment",],
    confint(lm_uncategorized)["treatment",]))

colnames(effect_sum)[3:4]<-c("confmin", "confmax")

effect_sum<-cbind(effect_sum,
                  100*effect_sum[,c("treatment_effect","confmin", "confmax")]/
                    filter(means, type=="charge")[c(6, 4, 1, 2, 3, 5),"control"])

colnames(effect_sum)[5:7]<-c("percent_treatment_effect", "percent_confmin", "percent_confmax")  
kable(effect_sum, caption="Effect estimates and confidence intervals")
```





```{r}
effect_plot<-ggplot()+
  geom_errorbar(data=effect_sum, mapping=aes(x=charge_type, ymin=percent_confmin, ymax=percent_confmax))+
  geom_point(data=effect_sum, aes(x=charge_type, y=percent_treatment_effect))+
  scale_y_continuous(breaks=seq(-60,40,10))+
  geom_point(data=charge_diffs, aes(x=group, y=percent_diff_in_means), color="blue")

effect_plot

```

In summary, our mean effect estimates are as follows: 

* ***Violent offenses:*** a ```r round(abs(effect_sum$percent_treatment_effect[1]),1)```% decrease in charges.

* ***Property offenses:*** a `r round(abs(effect_sum$percent_treatment_effect[2]),1)`% decrease in charges.

* ***Drug offenses:*** a `r round(abs(effect_sum$percent_treatment_effect[3]),1)`% decrease in charges.

* ***Firearms offenses:*** a `r round(abs(effect_sum$percent_treatment_effect[4]),1)`% increase in charges.

* ***Other offenses:*** a `r round(abs(effect_sum$percent_treatment_effect[5]),1)`% decrease in charges.

* ***Uncategorized offenses:*** a `r round(abs(effect_sum$percent_treatment_effect[6]),1)`% decrease in charges.

In all six categories, we have statistically significant evidence of a nonzero treatment effect (a decrease in all categories except firearms offenses). 

Comparing to the point estimates of differences in means (prior to controlling for arrest counts), we observe that none of these estimates fall within our 95% confidence intervals. For violent, property, drug, and other offenses, we find that our estimated effect is a smaller decrease than the percent change in means. In the case of uncategorized offenses, our estimated effect is a larger decrease than the difference in means. Finally, in the case of firearm offenses, the effect we estimate is that of a substantially smaller increase (about 8%) than the simple difference in means (about 24%).


## Discussion

An analysis of Krasner's office's effect on charges that does not take into account numbers of arrests is open to substantial confounding. We find that after controlling for arrest counts, the effect of being charged by Krasner's office was a decrease in charges in every category of offense except for firearms. For firearms offenses, we find that after accounting for arrests, which saw during Krasner's tenure as DA a `r round(abs(means[means$group=="firearms" & means$type=="arrest",]$percent_diff_in_means),1)`% increase from the pre-Krasner mean, Krasner's office had the effect of a `r round(abs(effect_sum$percent_treatment_effect[4]),1)`% increase in charges. 

Next, we discuss potential limitations of our study and examine how robust our study is to violations of our assumptions and model choices through some sensitivity analyses.

We recall that our estimates are not the entirety of Krasner's DAO's effect on charges. For instance, a decision early in Krasner's tenure to drop all marijuana possession charges may well have had an effect on PPD policy on whether to make arrests in cases of marijuana possession. Thus, by its effect on arrests, Krasner's DAO may have had an additional effect on the reduction in charges for drug offenses. However, any such effect is not included in our estimate--the above estimates only apply to what effect Krasner's office had in charging the arrests they were given as compared to Williams' DAO.

Our assumption that there was only one form of the treatment is almost certainly not entirely accurate: changes to charging policy through Krasner's tenure may well mean that the same day being charged by Krasner's DAO at one time could have different outcomes from being charged during a different time. In particular, Krasner's first several months as DA involved changes to the DAO, its policies, and its personnel, including large turnover in staff and prosecutors. We re-run our analysis, excluding the first 3 months of Krasner's tenure to see if it changes our conclusions:

```{r}
charges_res<-charges_all[charges_all$date_value<="2017-06-29" |
                              charges_all$date_value>="2018-03-01",]
n_treatment<-nrow(charges_res[charges_res$treatment,])
n_control<-nrow(charges_all)-n_treatment

m_res.out<-matchit(treatment~arrests_violent+arrests_property+arrests_drugs+arrests_other+arrests_firearms+arrests_uncategorized, data=charges_res, method="nearest", caliper=0.1, ratio=1.7, min.controls=1, max.controls=3)

matched_data_res<-match.data(m_res.out)
matched_data_res$treatment<-as.numeric(matched_data_res$treatment)


lm_violent_res.out<-lm(charges_violent~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data_res)
lm_property_res.out<-lm(charges_property~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data_res)
lm_drugs_res.out<-lm(charges_drugs~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data_res)
lm_firearms_res.out<-lm(charges_firearms~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data_res)
lm_other_res.out<-lm(charges_other~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data_res)
lm_uncategorized_res.out<-lm(charges_uncategorized~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data_res)

effect_sum[,"treatment_effect_res"]<-
  rbind(lm_violent_res.out$coefficients["treatment"],
    lm_property_res.out$coefficients["treatment"],
    lm_drugs_res.out$coefficients["treatment"],
    lm_firearms_res.out$coefficients["treatment"],
    lm_other_res.out$coefficients["treatment"],
    lm_uncategorized_res.out$coefficients["treatment"])
                                
effect_sum<-cbind(effect_sum,
  rbind(confint(lm_violent_res.out)["treatment",],
    confint(lm_property_res.out)["treatment",],
    confint(lm_drugs_res.out)["treatment",],
    confint(lm_firearms_res.out)["treatment",],
    confint(lm_other_res.out)["treatment",],
    confint(lm_uncategorized_res.out)["treatment",]))

colnames(effect_sum)[9:10]<-c("confmin_res", "confmax_res")

select(effect_sum, "treatment_effect", "confmin", "confmax", "treatment_effect_res", "confmin_res", "confmax_res")

```

While we see some small changes in our estimates, our confidence intervals still do not contain zero, and we come to the same broad conclusions as before. 

Next, we consider possible bias in our estimates introduced by our matching procedure, particularly the fact that we discarded treatment units in forming our matched data. First, we note that achieving good balance on covariates reduces interpolation bias (as defined in @KZ). Since we are calculating the average treatment effect (ATE) and not the average treatment effect on the treated (ATT), we not need to worry that dropping treated units changes the population for which we make our estimates. However, by dropping both treatment and control units, there is potential that we introduce extrapoation bias into the study. 

First, we argue this is not a concern in our study. The reason is that we assume that the DAO's decisions to charge individual cases are independent of the volume or distribution of the other cases (arrests) from that day. Consequently, we argue that we can extrapolate effects computed from days with a very distribution of types and counts or arrests to days with very different distributions and counts of arrests. Consequently, discarding some of our treatment units with propensity scores that are harder to match should not introduce substantial extrapolation bias as the units were discarded on the basis of arrest covariates (via propensity scores)  and not on the basis of outcomes.

However, to check that our analysis is robust to changes in the matching mechanism, we report the results of several other matching methods that were employed. The code for these is contained in the appendix.

To assess our assumption of ignorability, we employ a sensitivity analysis method developed by Blackwell @Blackwell. We measure sensitivity to one-sided selection bias: that is, the amount of an unmeasured confounder could increase (or decrease) charge counts, controlling for arrests, in the treatment group (cases post 2018) compared to the control group (pre-2018). Such confounding would exist if, for instance, quality of police reports or prior records of arrestees changed between or during the two DA administrations, making arrestees more or less susceptible to charges. Following Blackwell's method, for each amount of confounding (alpha), we assess whether this would reverse our conclusion that having Krasner's DAO charge cases results in a statistically significant change in number of charge in each category, controlling for arrests. We report the amount of confounding (measured in the original units of numbers of charges) that would be needed to reverse our conclusion.


```{r sensitivity}


model.t<-glm(treatment~arrests_violent+arrests_property+
               arrests_drugs+arrests_firearms+arrests_other+
               arrests_uncategorized, data=matched_data2, 
             family=binomial())

alpha_violent_1=seq(-8, 2, by=0.01)
sens_violent_1 <- causalsens(lm_violent, model.t, 
                      ~ arrests_violent+arrests_property+arrests_drugs+
                        arrests_firearms+arrests_other+arrests_uncategorized,
                      data = matched_data2,  confound = one.sided, alpha=alpha_violent_1)
alpha_property_1=seq(-6, 0, by=0.01)
sens_property_1 <- causalsens(lm_property, model.t, 
                      ~ arrests_violent+arrests_property+arrests_drugs+
                        arrests_firearms+arrests_other+arrests_uncategorized,
                      data = matched_data2,  confound = one.sided, alpha=alpha_property_1)
alpha_drugs_1=seq(-10, 0, by=0.01)
sens_drugs_1<-causalsens(lm_drugs, model.t, 
                      ~ arrests_violent+arrests_property+arrests_drugs+
                        arrests_firearms+arrests_other+arrests_uncategorized,
                      data = matched_data2,  confound = one.sided, alpha=alpha_drugs_1)
alpha_firearms_1=seq(-0.5, 0.5, by=0.01)
sens_firearms_1<-causalsens(lm_firearms, model.t, 
                      ~ arrests_violent+arrests_property+arrests_drugs+
                        arrests_firearms+arrests_other+arrests_uncategorized,
                      data = matched_data2,  confound = one.sided, alpha=alpha_firearms_1)
alpha_other_1=seq(-4, 0, by=0.01)
sens_other_1<-causalsens(lm_other, model.t, 
                      ~ arrests_violent+arrests_property+arrests_drugs+
                        arrests_firearms+arrests_other+arrests_uncategorized,
                      data = matched_data2,  confound = one.sided, alpha=alpha_other_1)
alpha_uncategorized_1=seq(-4, 0, by=0.01)
sens_uncategorized_1<-causalsens(lm_uncategorized, model.t, 
                      ~ arrests_violent+arrests_property+arrests_drugs+
                        arrests_firearms+arrests_other+arrests_uncategorized,
                      data = matched_data2,  confound = one.sided, alpha=alpha_uncategorized_1)

#a function to extract the value of alpha (amoung of selection bias=amount of confounding) 
#and the R-squared at which our causal estimates would lose statistical significance.
cnf_val<-function(X)
{
  ind<-min(which(X$lower<0))
  c(X[ind, "alpha"], X[ind, "rsqs"])
}
confounding_values<-as.data.frame(
  rbind(
    cnf_val(sens_violent_1$sens),
    cnf_val(sens_property_1$sens),
    cnf_val(sens_drugs_1$sens),
    cnf_val(sens_firearms_1$sens),
    cnf_val(sens_other_1$sens),
    cnf_val(sens_uncategorized_1$sens)
  )
)

types<-c("violent", "property", "drugs", "firearms", "other", "uncategorized")
row.names(confounding_values)<-types
colnames(confounding_values)<-c("alpha", "rsqs")

control_means<-filter(means, type=="charge")[, c("group", "control")][c(6, 4, 1, 2, 3, 5),]
confounding_values[,"percent"]<-confounding_values$alpha/control_means$control



par(mfrow=c(2, 3))
plot(sens_violent_1, type = "raw", bty = "n", main="Violent (one-sided)")
plot(sens_property_1, type = "raw", bty = "n", main="Property (one-sided)")
plot(sens_drugs_1, type = "raw", bty = "n", main="Drugs (one-sided)")
plot(sens_firearms_1, type = "raw", bty = "n", main="Firearms (one-sided)")
plot(sens_other_1, type = "raw", bty = "n", main="Other (one-sided)")
plot(sens_uncategorized_1, type = "raw", bty = "n", main="Uncategorized (one-sided)")



```

The results of our sensitivity analysis are as follows:

* ***Violent offenses:*** An unobserved covariate would have to on average result in `r abs(confounding_values["violent", "alpha"])` (a `r round(abs(confounding_values["violent", "percent"])*100,1)` decrease from control) fewer charges per day between cases tried pre and post Krasner to reverse our conclusion that Krasner's DAO caused a statistically significant decrease in violent charges controlling for arrests.

* ***Property offenses:*** An unobserved covariate would have to on average result in `r abs(confounding_values["property", "alpha"])` (a `r round(abs(confounding_values["property", "percent"])*100,1)` decrease from control) fewer charges per day between cases tried pre and post Krasner to reverse our conclusion that Krasner's DAO caused a statistically significant decrease in property charges controlling for arrests.

* ***Drug offenses:*** An unobserved covariate would have to on average result in `r abs(confounding_values["drugs", "alpha"])` (a `r round(abs(confounding_values["drugs", "percent"])*100,1)` decrease from control) fewer charges per day between cases tried pre and post Krasner to reverse our conclusion that Krasner's DAO caused a statistically significant decrease in drug charges controlling for arrests.

* ***Firearms offenses:*** An unobserved covariate would have to on average result in `r abs(confounding_values["firearms", "alpha"])` (a `r round(abs(confounding_values["firearms", "percent"])*100,1)` increase from control) more charges per day between cases tried pre and post Krasner to reverse our conclusion that Krasner's DAO caused a statistically significant increase in firearms charges controlling for arrests.

* ***Other offenses:*** An unobserved covariate would have to on average result in `r abs(confounding_values["other", "alpha"])` (a `r round(abs(confounding_values["other", "percent"])*100,1)` decrease from control) fewer charges per day between cases tried pre and post Krasner to reverse our conclusion that Krasner's DAO caused a statistically significant decrease in other charges controlling for arrests.

* ***Uncategorized offenses:*** An unobserved covariate would have to on average result in `r abs(confounding_values["uncategorized", "alpha"])` (a `r round(abs(confounding_values["uncategorized", "percent"]) *100,1)` decrease from control) fewer charges per day between cases tried pre and post Krasner to reverse our conclusion that Krasner's DAO caused a statistically significant decrease in uncategorized charges controlling for arrests.

Based on our sensitivity analysis, we conclude that our treatment effect estimates for property, other, and uncategorized offenses are robust to effects of potential unobserved covariates and our conclusions for drug and violent offenses are moderately so. However, our conclusion that Krasner's office caused an increase in charges for firearms offenses controlling for arrests is quite sensitive to the effects of potential unobserved covariates. 

In conclusion, our analysis shows that Krasner's DAO had a statistically significant effect on the number of cases charged for each offense type, a decrease in 5 of 6 categories. None of these estimates are as large in magnitude as simple point estimates (prior to controlling for arrests) would suggest. The largest estimated effect we observe is a 34.8% decrease in charges for other offenses (sex work related offenses and threats of violence), followed by a 15.6 percent decrease in charges for property related offenses. For firearms offenses, despite the fact that arrests increased steeply during roughly the same time frame as charges, we find that the effect of Krasner's DAO was an 8% increase in charges. However, our sensitivity analysis suggests that an unobserved covariate would only have to be responsible for a small change in charge counts to reverse this conclusion. 

## Future Directions

Our sensitivity analysis above suggests that the extent to which our ignorability assumption holds could reverse our conclusion, particularly that Krasner's DAO was responsible for an increase in firearms charges. If we had more data on potential factors that effect charges such as prior criminal records of defendants, changes to state laws and policies regarding how arrests are charged, and quality of police reports and data provided to the DAO, we would be better able to evaluate this assumption.

The longitudinal nature of our data also suggests that a model that takes this into account could provide more accurate effect estimates. For instance, an interrupted time series model could be well suited for this problem. One challenge is the lack of a control population post 2018 (after Krasner took office). Future work in this direction could proceed by finding charge and arrest data from a comparison population in a different city which did not experience a change in DA during our time-frame and use this in an interrupted time series framework. 

Finally, beyond charges, the DAO plays a substantial role later in the criminal justice process, particularly in the bail process, diversionary programs, and sentencing. It would be interesting to implement some of these methods (controlling for the actions of other actors, such as judges, in the process) to estimate the effects of Krasner's DAO in these areas. Such studies would create a more comprehensive picture of Krasner's DAO's effect on the criminal justice system in Philadelphia.

## Bibliography

