---
title: "Matching"
author: "Vitaly Lorman"
date: "3/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(knitr)
library(lubridate)
library(scales)
library(reshape2)
library(MatchIt)
library(optmatch)
library(Zelig)
```

In this document, we keep track of our various attempts at matching and analysis of balance.

```{r read data}
charges_all<-read.csv("charges_all.csv", row.names=1)
charges_all_long<-read.csv("charges_all_long.csv", row.names=1)
charges_all$date_value<-as.Date(charges_all$date_value)
charges_all_long$date_value<-as.Date(charges_all_long$date_value)
```


### Nearest neighbor matching
```{r match 1nearest neighbors}
##default
m_nearest.out<-matchit(treatment~arrests_violent+arrests_property+arrests_drugs+arrests_other+arrests_firearms+arrests_uncategorized, data=charges_all, method="nearest")
s_nearest.out <- summary(m_1nearest.out, standardize = TRUE)
s_nearest.out
plot(s_nearest.out)

plot(m_1nearest.out,  type = "jitter", interactive = FALSE)
plot(m_1nearest.out,  type = "hist")
plot(m_1nearest.out, type="qq")

```


```{r match knearest}
m_knearest.out<-m_nearest.out<-matchit(treatment~arrests_violent+arrests_property+arrests_drugs+arrests_other+arrests_firearms+arrests_uncategorized, data=charges_all, method="nearest", min.controls=1, max.controls=3, ratio=1275/653)
s_nearest.out <- summary(m_1nearest.out, standardize = TRUE)
  
s_knearest.out <- summary(m_knearest.out, standardize = TRUE)
s_knearest.out

plot(s_knearest.out)
plot(m_knearest.out,  type = "jitter", interactive = FALSE)
plot(m_knearest.out,  type = "hist")
plot(m_knearest.out, type="qq")
```

Does not seem to do better.

```{r matching subclass}
m_subclass.out<-matchit(treatment~arrests_violent+arrests_property+arrests_drugs+arrests_other+arrests_firearms+arrests_uncategorized, data=charges_all, method="subclass")
s_subclass.out <- summary(m_subclass.out, standardize = TRUE)
s_subclass.out
plot(s_subclass.out)

plot(m_subclass.out,  type = "jitter", interactive = FALSE)
plot(m_subclass.out,  type = "hist")
plot(m_subclass.out, type="qq")
```

Subclass matching seems to do really well.

```{r matching optimal}
m_optimal.out<-matchit(treatment~arrests_violent+arrests_property+arrests_drugs+arrests_other+arrests_firearms+arrests_uncategorized, data=charges_all, method="optimal", ratio=1275/653)
s_optimal.out <- summary(m_optimal.out, standardize = TRUE)
s_optimal.out
plot(s_optimal.out)

plot(m_optimal.out,  type = "jitter", interactive = FALSE)
plot(m_optimal.out,  type = "hist")
plot(m_optimal.out, type="qq")


```
Nope skip this one!


```{r matching cem}

m_cem.out<-matchit(treatment~arrests_violent+arrests_property+arrests_drugs+arrests_other+arrests_firearms+arrests_uncategorized, data=charges_all, method="cem")
s_cem.out <- summary(m_cem.out, standardize = TRUE)
s_cem.out
plot(s_cem.out)

plot(m_cem.out,  type = "jitter", interactive = FALSE)
plot(m_cem.out,  type = "hist")
plot(m_cem.out, type="qq")




```

CEM drops too many observations

Try discarding everything outside the convex hull

```{r discard hull}
m_hull.out<-matchit(treatment~arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_uncategorized, data=charges_all, method="nearest", discard="both")

s_hull.out <- summary(m_hull.out, standardize = TRUE)
s_hull.out
plot(s_hull.out)

plot(m_hull.out,  type = "jitter", interactive = FALSE)
plot(m_hull.out,  type = "hist")
plot(m_hull.out, type="qq")

m_hull_subclass.out<-matchit(treatment~arrests_violent+arrests_property+arrests_drugs+arrests_other+arrests_firearms+arrests_uncategorized, data=charges_all, method="subclass", discard="both")
s_hull_subclass.out<-summary(m_hull_subclass.out, standardize=TRUE)
s_hull_subclass.out

plot(s_hull_subclass.out)
plot(m_hull_subclass.out,  type = "jitter", interactive = FALSE)
plot(m_hull_subclass.out,  type = "hist")
plot(m_hull_subclass.out, type="qq")

```

Try nearest neighbor with calipers (ok to drop some treatment units)

```{r calipers}
mcal1.out<-matchit(treatment~arrests_violent+arrests_property+arrests_drugs+arrests_other+arrests_firearms+arrests_uncategorized, data=charges_all, method="nearest", caliper=0.01)
scal1.out <- summary(mcal1.out, standardize = TRUE)
scal1.out
plot(scal1.out)

plot(mcal1.out,  type = "jitter", interactive = FALSE)
plot(mcal1.out,  type = "hist")
plot(mcal1.out, type="qq")



```
May still need to handle other manually.

## Matching with subclasses

The number of arrests for various offenses is the biggest potential confounder of charges. We account for this by matching. After experimenting with different methods of matching, we found the best balance using subclass matching while discarding a handful of observations (?) outside the common support of the propensity scores. This is justified because...

```{r matching}
m.out<-matchit(treatment~arrests_violent+arrests_property+arrests_drugs+arrests_firearms++arrests_other+arrests_uncategorized, data=charges_all, method="subclass", discard="both")

s.out <- summary(m.out, standardize = TRUE)
s.out

plot(s.out)
plot(m.out,  type = "jitter", interactive = FALSE)
plot(m.out,  type = "hist")
plot(m.out, type="qq", interactive=FALSE, subclass=1:6)

matched_data<-match.data(m.out)
summarise(group_by(matched_data, subclass, treatment), count=n())

```

Let's look at the distributions of arrests in the different classes:
```{r}
matched_long<-melt(matched_data, id=c("date_value", "subclass", "treatment", "weights", "distance"))

subclass_sum<-matched_long %>%
  group_by(subclass, variable, treatment) %>%
  summarise(mean=mean(value))

subclass_arrest_sum<-subclass_sum[grep("^arrests", subclass_sum$variable),]

ggplot(data=subclass_arrest_sum, aes(x=variable, y=mean, fill=treatment))+
  geom_bar(position="dodge", stat="identity")+
  facet_wrap(~subclass)+
  theme(axis.text.x = element_text(angle = 90))


```

## Using linear regression to estimate causal effects

We now use linear regression to estimate causal effects on the matched data. We carry this out by subclass, then aggregate the results.

```{r reg, echo=FALSE}
z_violent.out<-zelig(charges_violent~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data, mode="ls", by="subclass")
#x.out <- setx(z.out, data = match.data(m.out), fn = NULL, cond = TRUE)
#s.out <- sim(z.out, x = x.out)
s_violent<-summary(z_violent.out)

extract_coef<-function(coef_list){
  coef_list["treatmentTRUE"]
}
violent_coefs<-as.numeric(lapply(coef(z_violent.out), extract_coef))

colnames(matched_data)

violent_subclass_control_means<-matched_data %>%
  select(date_value, charges_violent, treatment, subclass)%>%
  filter(!treatment) %>%
  group_by(subclass) %>%
  summarise(mean=mean(charges_violent))

violent_subclass_control_means<-violent_subclass_control_means$mean

violent_coefs/violent_subclass_control_means

lm_violent_ag<-lm(charges_violent~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data, weights=weights)
summary(lm_violent_ag)



z_property.out<-zelig(charges_property~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data, mode="ls", by="subclass")
#x.out <- setx(z.out, data = match.data(m.out), fn = NULL, cond = TRUE)
#s.out <- sim(z.out, x = x.out)
summary(z_property.out)

lm_property_ag<-lm(charges_property~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data, weights=weights)
summary(lm_property_ag)

z_drugs.out<-zelig(charges_drugs~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data, mode="ls", by="subclass")
#x.out <- setx(z.out, data = match.data(m.out), fn = NULL, cond = TRUE)
#s.out <- sim(z.out, x = x.out)
summary(z_drugs.out)

lm_drugs_ag<-lm(charges_drugs~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data, weights=weights)
summary(lm_drugs_ag)

z_firearms.out<-zelig(charges_firearms~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data, mode="ls", by="subclass")
#x.out <- setx(z.out, data = match.data(m.out), fn = NULL, cond = TRUE)
#s.out <- sim(z.out, x = x.out)
summary(z_firearms.out)

lm_firearms_ag<-lm(charges_firearms~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data, weights=weights)
summary(lm_firearms_ag)

z_other.out<-zelig(charges_other~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data, mode="ls", by="subclass")
#x.out <- setx(z.out, data = match.data(m.out), fn = NULL, cond = TRUE)
#s.out <- sim(z.out, x = x.out)
summary(z_other.out)

lm_other_ag<-lm(charges_other~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data, weights=weights)
summary(lm_other_ag)

z_uncategorized.out<-zelig(charges_uncategorized~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data, mode="ls", by="subclass")
#x.out <- setx(z.out, data = match.data(m.out), fn = NULL, cond = TRUE)
#s.out <- sim(z.out, x = x.out)
summary(z_uncategorized.out)

lm_uncategorized_ag<-lm(charges_uncategorized~treatment+arrests_violent+arrests_property+arrests_drugs+arrests_firearms+arrests_other+arrests_uncategorized, data=matched_data,  weights=weights)
summary(lm_uncategorized_ag)


ag_coefs<-c(lm_violent_ag$coefficients["treatmentTRUE"], 
            lm_property_ag$coefficients["treatmentTRUE"],
            lm_drugs_ag$coefficients["treatmentTRUE"],
            lm_firearms_ag$coefficients["treatmentTRUE"],
            lm_other_ag$coefficients["treatmentTRUE"],
            lm_uncategorized_ag$coefficients["treatmentTRUE"])



```
